### 2015-02-07 icfp15-servo/abstract.tex
Abstract

%!TEX root = paper.tex
%
\begin{abstract}
Inlining is an optimization that replaces a call to a function with that function's 
body.
This optimization not only reduces the overhead of a function call, but can 
expose additional optimization opportunities to the compiler, such as removing
redundant operations or unused conditional branches.
Another optimization, copy propagation, replaces a redundant copy of a still-live 
variable with the original.
Copy propagation can reduce the total number of live variables, reducing register
pressure and memory usage, and possibly
eliminating redundant memory-to-memory copies.
In practice, both of these optimizations are implemented in nearly every
modern compiler.

These two optimizations are practical to implement and effective in first-order 
languages, but in languages with lexically-scoped first-class functions (aka, closures), these optimizations 
are not available to code programmed in a higher-order style.
With higher-order functions, the analysis challenge has been that the 
environment at the call site must be the same as at the closure capture location, 
up to the free variables, or the meaning of the program may change.
Olin Shivers' 1991 dissertation called this family of optimizations 
\emph{\superbeta} and he proposed one analysis technique, called \emph{reflow},
to support these optimizations.
Unfortunately, reflow has proven too expensive to implement in practice.
Because these higher-order optimizations are not available in functional-language
compilers, programmers studiously avoid uses of 
higher-order values that cannot be optimized (particularly in compiler
benchmarks).

This paper provides the first practical and effective technique for 
\superbeta (higher-order) inlining and copy propagation, which
we call unchanged variable analysis.
We show that this technique is practical by implementing it in the 
context of a real compiler for an ML-family language and showing 
that the required analyses have costs below 3\% of the total 
compilation time.
This technique's effectiveness is shown through a set of benchmarks
and example programs, where this analysis exposes additional potential
 optimization sites.
\end{abstract}

>>>>

%!TEX root = paper.tex
%
\begin{abstract}
The Servo project is an effort started at Mozilla Research to build a new web browser
engine that takes advantage of both modern parallel hardware and advances in language
technology that prevent many of the common security errors that plague all of the other
modern browsers.
The new language, Rust, enabled us to take on this project in a type-centric way instead of
the traditional C++ that relies on coding conventions, imprecise static analyses, and
fuzzers.
While many languages before have offered some combination of memory safety, parallelism,
and concurrency, Rust is, to our knowledge, the first to do so that provides the programmer
with the control over both memory and generated code that is required by systems programmers.
In this paper, \emph{systems} \emph{software} refers to programs that have some combination
of soft realtime requirements and memory, CPU, and power usage constraints that require the
programmer to be able to predict these features from their source code --- without being a
compiler developer.

Servo is currently over 300k lines of code and implements enough of the web to render and
process many pages, though it is still a far cry from the over 7 million lines of code in
the Mozilla Firefox browser and its associated libraries.
However, we believe that we have implemented enough of the web platform to provide an
early report on the successes, failures, and open problems remaining in Servo, from the
point of view of programming languages and runtime research.
It is our hope that this work will spur additional research in new language features
suitable for a systems language that also scale up to projects the size of a web browser
engine.
\end{abstract}
